### 2018-05-21 papers/adewale/acknowledgements.tex

I would like to thank Thomas Schaffner, Chris Galea, and Donna Gabai for helping and/ or supporting me in my Thesis Presentation. I would like to thank Aur\`ele Barri\`ere for co-implementing the B+-Tree library. He was very instrumental in the B+-Tree library's final implementation, which leverages the B+-Tree with cursor concept. I would also like to thank Professor Andrew Appel for advising me throughout this thesis project and Professor Margaret Martonosi for being my thesis reader. I would like to also thank my parents, Segun and Victoria Adewale, for their support through out my life and education. I doubt I would have made it this far without them. Finally, I am ultimately grateful to God for bringing me this far and for the opportunities that lie ahead. Soli Deo Gloria!