### 2020-07-13 papers/luke/poplsrc19/extended_abstract.tex

\documentclass[sigplan,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}

\acmConference[POPL SRC'19]{ACM SIGPLAN Symposium on Principles of Programming
  Languages -- Student Research Competition}{January 13--19,
  2019}{Cascais/Lisbon, Portugal} \acmYear{2019}
\acmISBN{}
\acmDOI{}
\startPage{1}

\bibliographystyle{ACM-Reference-Format}
\usepackage{booktabs}
\usepackage{subcaption}
\begin{document}

\title[]{Verification of a Cache-optimized Data Structure}
\author{Yixuan Chen} \affiliation{ \institution{University of Michigan} }
\email{xlk@umich.edu}

\author{Aur\`ele Barri\`ere} \affiliation{ \institution{\'Ecole Normale
    Sup\'erieure de Rennes} } \email{aurele.barriere@ens-rennes.fr}

\author{Lennart Beringer} \affiliation{ \institution{Princeton University} }
\email{eberinge@cs.princeton.edu}

\author{Andrew W. Appel} \affiliation{ \institution{Princeton University} }
\email{appel@cs.princeton.edu}

\begin{abstract}
  Recent years have witnessed a rapid development of main-memory database
  systems thanks to the growingly affordable memory. DeepSpecDB is a main-memory
  database management system implemented in C with deep specifications and
  end-to-end verification guaranteeing the correctness of the system. In this
  paper, we will present current status of DeepSpecDB on verification of index
  data structures.
\end{abstract}

\keywords{formal verification, deep specifications, database}

\maketitle

\section{Introduction}

As the unit price of memory decreases over time, efforts have been put into
migrating the traditional disk-based applications into main memory. A classical
category of such applications are databases. For example, H-store~\cite{hstore},
MongoDB, Redis and Memcached are well-known database systems featuring
main-memory index and storage.

Meanwhile, database systems are the kind of critical applications where security
vulnerabilities can lead to catastrophic consequences. We believe that formally
verifying functional correctness of programs with respect to deep specifications
is the remedy. Deep specifications are rich, live, formal, and two-sided,
ensuring that the behavior will be captured by the specifications and proved
correctly by machine-checked proofs. With the help of the Verified Software
Toolchain~\cite{VST}, we are now able to verify C programs against the
CompCert~\cite{compcert} operational semantics. The DeepSpecDB project aims to
design and verify a main-memory database system with deep specification
guaranteeing the correctness of the system. In this paper, we present our
efforts on the indexing data structures, which databases use to organize the
data and speed up data access.

\section{Overview of the data structure}

We design our database index based on MassTree~\cite{masstree}. In general, the
data structures maps from variable-length strings to values. MassTree maintains
the mapping by organizing strings in a \textit{trie over B+ trees}: each node of
traditional trie is now replaced with a B+ tree and a trie node now corresponds
to a fixed-length slice of key rather than a single character. To account for
the possible residual suffix of keys, a data structure named \textit{border
  node} is introduced. The \textit{border node} replaces the leaf node of
original B+ tree, and can potentially point to the next layer of trie nodes or
client values corresponding to different prefixes of current slice. The length
of the slice depends on the CPU word size, so that we can express the slice as
an integer and speed up the operations. The fanout of the B+ trees as well as
other parameters of MassTree are chosen so that internal structures can fit well
in cache line which improves the performance. Although the original MassTree was
designed for key-value databases, later development of Silo~\cite{Silo} and
SiloR~\cite{SiloR} suggest its capability of indexing for relational databases.

In the original MassTree design, the \textit{border node}s are linked with each
other to facilitate range queries and removals. We decide to remove the links in
favor of a new structure called \textit{cursor} inspired by SQLite. Abstractly,
if all keys are organized in an ordered list, a \textit{cursor} should point
between two adjacent keys and any key bounded by the two can be associated with
the \textit{cursor}. Concretely, the \textit{cursor} for B+ trees is implemented
with a list of pointers pointing at various internal nodes and the
\textit{border node}, and the indexes into those nodes. The list should also be
the trace when one tries to access the \textit{border node} given any slice of
key associated with the cursor from the root node. The \textit{cursor} for a
trie is a list of B+ tree cursors, which is also the trace when one tries to
access the client value given any key associated with the cursor.

\textit{Cursor}s provide similar sequential performance of range queries to
links because one can access the next or previous record in the structure by
traversing upward and then downward with the help of \textit{cursor}s, which
should cost only amortized constant time. Moreover, the traversal should hit
cache frequently considering the locality of range queries. We favor
\textit{cursor}s over links in the data structures because we anticipate better
concurrent performance compared to implementation with links for two reasons.
First, \textit{cursor}s can be allocated as thread local objects leading to less
contentions. Second, we expect \textit{cursor}s to perform better when
overwriting update is not available (updating the links in this case is
expensive), which happens in the optimistic concurrency control variant from
Silo.

\section{Modular Coq-based verification using VST}

The verification proceeds in a modular approach. In the database, the B+ trees
are also used as the index for integer keys, and we build tries on exactly the
same interface exposed to the rest of database. We organize the verification in
the same structure as we write the program. The verification of B+
trees~\cite{aurele} abstracts the implementation into specification. Tries are
then verified given only the specification of B+ trees. In another sense, the
verification holds for any data structure that is verified with the same
specification. Furthermore, the specification for the database index is
parameterized by key types, which means the specification of tries is the same
as that of B+ trees modulo the theory of keys. This also enables us to use the
verification of trie as a test for the specification: only if the specification
is neither over-restricted nor under-specified can the proofs go through.

In another direction, the application domain proof is modularized away from the
concrete semantics in verification of both data structures. That is, C programs
are proved to correctly implement corresponding functional programs. The
refinement is established in Verifiable C using Verified Software Toolchain. The
functional programs are proved to conform to the specification, which is usually
properties about keys, values and interaction with the data structures. The
benefit of modularization is significant: we need not care about the low level
details, such as pointers and memory management when reasoning about functional
correctness. We can also forget the application logic when we are in the
separation logic world.

\section{Lessons learned}

During the verification, we also discover several methodologies that is useful
for verification of complex data structures, especially with support of
\textit{cursor}s.

\subsection{Augmented types}

When using separation logic, one needs to derive the abstract data types
corresponding to the concrete ones, where all the addresses are typically
abstracted away (for example, in the proof for list reversal~\cite{reynolds}).
However, a \textit{cursor} might point to internal nodes of a data structure
which is often achieved by saving copies of pointers in C. We find it hard to
express the same relation using abstract cursor and the abstract data structure.
One promising solution is to use magic wand~\cite{wand}, but this approach turns
out to be unsound. Therefore we introduce the augmented types, where we include
the addresses in the abstract data type in order to express the relation between
\textit{cursor} and the data structure without violating the soundness.

Augmented types, however, introduce another problem: the addresses of data
structures are usually determined by allocation at runtime and data structures
located in different address space can express the same abstract data. To
account for the non-determinism, we type our specifications for mutating
operations as predicates rather than functions. It would be ideal if we can
still justify the determinism excluding the addresses, which we will see in the
next section.

\subsection{Flattening}

Coq has a very strict type system to ensure the strict positiveness of
inductive types and termination of recursive functions. It ensures the soundness
of the system, but also adds to the difficulty to define \textit{trie over B+
  trees}: with B+ trees hidden from trie's definition, Coq cannot check the strict
positiveness and rejects the definition. We introduce the concept of
\textit{flattening} to pass the type check. In short, an extra function
\textit{flatten} is required in the specification which turns an abstract data
structure into an ordered list of key-value pairs containing the same set of
data. We can then define the data structure for trie based on \textit{flattened}
B+ trees, which can pass Coq's type check without problem.

\textit{Flattening} excels among other candidate solutions to the above problem
because it synergizes with other components of verification well. We find
similar idea in Tobias Nipkow's verification of search trees~\cite{nipkow},
where \textit{flattening} is used to verify the functional correctness. We find
it also helpful in our verification: many theorems would have been proved
separately for different data structures can now be proved once for the
\textit{flattened} list and work for all. \textit{Flattening} also implies the
determinism of data structure mutations: if the mutating operations commute with
the \textit{flatten} function, we are guaranteed the determinism by the
orderedness of the \textit{flattened} list.

\section{Current status and the future}

DeepSpecDB is still an ongoing project
(\url{https://github.com/PrincetonUniversity/DeepSpecDB}). We have almost
finished the verification for B+ trees, and basic operations for tries. However,
the current verification targets only sequential version of the database. To
achieve high performance, we are going to rewrite the program and the
verification to support concurrency. We will also benchmark different
implementations to confirm our expectation of the performance. Furthermore,
there are efforts to integrate the verification of index engines with other
components of database, such as query compilers. We expect an end-to-end
verification will be established from the semantics of SQL to the semantics of
assembly code.

%% Bibliography
\bibliography{../reference.bib}

\end{document}